---
title: "2_Simulating_Measurement_Error"
output: html_document
date: "2024-12-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
library(ggpubr)
```

## Data set of Simulation

Based on the manual of the ACT-measuring device used in the study, we simulated a measurement error with coeficient of variation of 10%. Based on the MCMC of the [rethinking package of RMcElreath.](https://github.com/rmcelreath/rethinking), we simulated the distribution of P(ACT\|kACT). (= the distribution of the real ACT for a measured value kACT). From this distribution we took 1000 random samples.

```{r}
sampling_kact <- function(kact,nsamples){
  d <- list(kact= kact
            )
  model1 <- ulam(
    alist(
      kact ~ dnorm(act_bar,act_bar/10),
      
      #prior
      act_bar ~ dunif(0,1000)
    ), data= d, chains =4, cores=4, iter=10000
  )
  samples_pre <- rethinking::extract.samples(model1, n=nsamples)
  head(samples_pre[[1]],nsamples)
  
} 



```

A data set was build for the following situation:\

| Parameter | Explanation                                                                                          |
|-------------|----------------------------------------------------------|
| act_0     | distribution of ACT before heparin administration when the measured value (kACT) = 116               |
| act_33000 | distribution of ACT after administration of 33000 IU of heparin when the measured value (kACT) = 535 |
| slope_hep | calculated heparin sensitivity for each sampled pair of act\_ and act_33000                          |
| act_x     | distribution of ACT for a measured ACT post protamin administration of 124                           |
| hep_act_x | calculated heparin level for each combination of act0,act_33000 and act_x sample.                    |

```{r}
set.seed(123)

dat1000<- tibble(
  act_0 = sampling_kact(116,1000), # 1000 samples from P(ACT|kACT=116)
  act_33000 = sampling_kact(535,1000), # 1000 samples from P(ACT|kACT=535)
  slope_hep= (act_33000-act_0)/33000, # resulting heparin sensitivity for each simulated pair of samples.
  act_x = sampling_kact(124,1000), # 1000 samples from P(ACT|kACT=124)
  hep_act_x = (act_x-act_0)/slope_hep # 1000 samples from P(Heparin| kACT=124)
  
)
```

Vizualisation of the distribution of possible (real)ACTs when a value of kACT of 116 is measured. (fig_C_extra)

```{r}
fig_C_extra <- ggplot(dat1000, aes(x=act_0))+
  geom_histogram(col='white',fill='steelblue', 
                 binwidth=5)+
  theme_classic()+
  labs(title='P(ACT|kACT=116)',
       subtitle = 'distribution of 1000 samples',
       x='ACT (seconds)')

fig_C_extra
```

## Fig C: P(kACT\|Heparin)

### heparine sensitivity

Repeating the procedure to calculate the slope/Heparin Sensitivity for each pair of samples from P(ACT\|kACT=116) and P(ACT\|kACT=535) gives fig C.
(Distribution of the ACT values pre and post Heparine administration are depicted as seperate boxplots.)

```{r}
fig_C <- ggplot(dat1000)+
    ## x-axis
  geom_hline(yintercept=0)+
    
  ## 1000 sampled sensitivity lines
  geom_segment(#data=.%>% head(150),
    aes(x= 0, xend=33000, y= act_0, yend=act_33000), col='grey70', 
    linewidth=0.5, alpha=0.1)+
  
  ## Box plot of P(ACT|kACT = 113)
  geom_boxplot(aes(y= act_0, x= -700),
               width=700, outliers = F,col='steelblue')+
  
  
  ## Boxplot of P(ACT|kACT= 535)
  geom_boxplot(aes(y= act_33000, x= 33000+700),width=700, outliers = T,
               col='steelblue')+
  
  
  
  
  ## y-axis
  geom_vline(xintercept=0)+
  
  theme_minimal()+
  labs(x='Heparin (IU)',
       y='Activated Clotting Time (s)')+
  
  ## measured sensitivity line
  annotate( geom='segment', x= 0, xend=33000, y= median(dat1000$act_0), yend=median(dat1000$act_33000), col='grey30', linewidth=2)+
  
  labs(title= 'P(ACT|heparin)',
       subtitle = 'Spread of Heparine Sensitivity based on 1000 samples')

fig_C

```

### P(ACT\|heparin = 20000): Predicting in the vertical plane

This simulation can also be used to predict what the real ACT would be if e.g. 20000 Heparin was given. This can be vizualized as the distribution of het intersection between horizontal line for Heparin=20000 and the 1000 simulated heparine sensitivity lines.

```{r}

plot_hep20000 <- ggarrange(fig_C + geom_vline(xintercept = 20000, col='red')+ labs(title='',subtitle=''),

dat1000 |> mutate(act_20000 = act_0+ slope_hep*20000) |>
  ggplot(aes(x= act_20000))+
  geom_histogram(color='white', fill='red', binwidth = 15)+
  theme_classic()+
  labs(x= 'ACT (s)')
)

fig_C_extra2 <-annotate_figure(plot_hep20000, top= text_grob("Simulated P(ACT|heparin=20000)"))
fig_C_extra2
```

## Fig D: P(heparin|kaACT): Predicting heparin level based on a measured ACT value 

Repeating the procedure to calculate the heparin level for each possible ACT value can be vizualized in fig D
```{r}
fig_D1 <- ggplot(dat1000)+
    ## x-axis
  geom_hline(yintercept=0)+
    
  ## 1000 sampled sensitivity lines
  geom_segment(#data=.%>% head(150),
    aes(x= 0, xend=33000, y= act_0, yend=act_33000), col='grey70', 
    linewidth=0.5, alpha=0.1)+
  
  ## Box plot of P(ACT|kACT = 113)
  geom_boxplot(aes(y= act_0, x= -500),
               width=700, outliers = F,col='steelblue')+
  
  
  ## Boxplot of P(ACT|kACT= 124)
  geom_boxplot(aes(y= act_x, x= -1000),width=700, outliers = T,
               col='darkred')+
  
  ## Intersection of each sample with the hep sensitivity slope --> Hep
  geom_point(aes(x=ifelse(hep_act_x<0,0, hep_act_x), y= act_x),size=0.1, col='darkred')+
  
  ## Distribution of hep_act_x
  geom_histogram(aes(x=ifelse(hep_act_x<0,0, hep_act_x), y=0.2*after_stat(count)),
                 binwidth=372, center= 0,fill='darkred', col='white',alpha=1)+
  
  ## y-axis
  geom_vline(xintercept=0)+
  
  theme_minimal()+
  labs(x='Heparin (IU)',
       y='Activated Clotting Time (s)')+
  
  ## measured sensitivity line
  annotate( geom='segment', x= 0, xend=33000, y= median(dat1000$act_0), yend=median(dat1000$act_33000), col='grey30', linewidth=2)+
  
  labs(title= 'P(heparin|kACT)',
       subtitle = 'Spread of Heparine levels based on 1000 samples')+
  ## Zooming in te region of interest
  
  coord_cartesian(xlim= c(-1000,7000), ylim= c(0,200)) +
  scale_x_continuous(minor_breaks = c(-1250,1250,3750,7250),
                     breaks = c(0,round(median(dat1000$hep_act_x),0),2500,5000),
                     labels=c(0,round(median(dat1000$hep_act_x),0),2500,5000))

fig_D1
```
Remark: AS negative Heparin can't exist, all negative calculated values are binned to Zero. We shoul bare in mind that these values are potentially associated with protamine overdose. (A more sophesticated simulation to incorporate this and the effect of protamine overdose on measured ACT value is would be needed.)

The median heparin is level 738 (corresponding to the value in fig B). This histogram can be used to calculate the change of Heparin level > 930 IU (see article)

```{r}
fig_D2 <- ggplot(dat1000)+
    ## x-axis
  geom_hline(yintercept=0)+
    
 
  ## Distribution of hep_act_x
  geom_histogram(aes(x=ifelse(hep_act_x<0,0, hep_act_x), y=after_stat(count)),
                 binwidth=372, center= 0,fill='darkred', col='white',alpha=0.7)+
  
  # median hep_act_x
  geom_vline(xintercept = median(dat1000$hep_act_x))+
  
  # Distribution of hep_act_X > 9300
  geom_histogram(data= dat1000 |> filter(hep_act_x > 930),
                 aes(x=ifelse(hep_act_x<0,0, hep_act_x), y=after_stat(count)),
                 binwidth=372, center= 0,fill='darkred', col='white',alpha=1)+
  
  
  annotate(geom = 'segment', x= 930, xend=930,y=0, yend= 150,size=1.5)+
  annotate(geom= 'segment', x= 930, xend=1400,y=150, yend= 150,size=1.5, 
           arrow=arrow(length=unit(3, "mm")), lineend='round')+
  annotate(geom='text', x = 3300, y = 150, 
           label= "P(Heparin >930|kACT=124) = 41%") +
  
  ## x-axis
  geom_hline(yintercept=0)+
  
  theme_minimal()+
 
  
  labs(x='Heparin (IU)',
    title= 'P(heparin|kACT)',
       subtitle = 'Spread of Heparine levels based on 1000 samples')+
  ## Zooming in te region of interest
  
  coord_cartesian(xlim= c(-1000,7000), ylim= c(0,300)) +
  scale_x_continuous(minor_breaks = c(-1250,1250,3750,7250),
                     breaks = c(0,round(median(dat1000$hep_act_x),0),2500,5000),
                     labels=c(0,round(median(dat1000$hep_act_x),0),2500,5000))
fig_D2
```

